Smart Representations: {R}ationality and Evolution in a Richer Environment

### 2016-01-29 paper/paperBlinded.tex
PHoS version

\documentclass[fleqn,reqno,11pt]{article}

@@

\usepackage{bigints}
\usepackage{MnSymbol}
\usepackage{a4wide}

@@

\title{Smart Representations: {R}ationality and Evolution in a Richer  Environment}
\author{Paolo Galeazzi \& Michael Franke} \date{current version: December 19, 2015}

@@

\begin{abstract}
  Standard applications of evolutionary game theory look at a single, fixed game and focus on
  the evolution of behavior for that game alone. Instead, this paper uses tools from
  evolutionary game theory to study the evolutionary competition between \emph{choice
    mechanisms} in a rich and variable environment. A choice mechanism is a way of representing
  a decision situation, paired with a method for choosing an act based on this 
  subjective representation. We demonstrate the usefulness of this approach by a case study
  that compares a number of classic proposals of what it means to make a rational choice. Our
  main result is that regret-based preference representations can be evolutionarily selected
  for, despite being non-veridical.
\end{abstract}

@@

If agents deal with a rich and variable environment, they have to face many different choice
situations.
Standard evolutionary game models frequently
simplify reality in two ways that are relevant for present purposes. Firstly, the environment
is represented as a fixed \emph{stage game}; secondly, the focus of
evolutionary selection is behavior for that stage game alone. In contrast, some argue for
studying the evolutionary competition of general \emph{choice mechanisms} in a rich and
variable environment
\citep[e.g.,][]{FawcettHamblin2013:Exposing-the-be,McNamara2013:Towards-a-Riche}. To show that
there is no tension here, this paper introduces a method for reconciling established notions of
evolutionary game theory with evolutionary selection of choice mechanisms in variable
environments \citep[see, inter alia, related work
by][]{Harley1981:Learning-the-Ev,ZollmanSmead2010:Plasticity-and-,SmeadZollman2013:The-Stability-o,OConnor2016:Evolving-to-Gen}.

@@

The subjective representation scheme $\gamma_i$ of agent $i$ is a function that
takes a game $G = \tuple{ N, (A_i , \pi_i)_{i \in N} }$ and maps it onto
$\gamma_i(G) = (\theta, e)$ where
$\theta \mycolon \bigtimes_{j \in N} A_j \rightarrow \mathbb{R}$ is a subjective utility function (the \emph{preference type} of the player), and $e$ is some subjective representation of a player's uncertainty about the co-players'
behavior (the \emph{belief type} of the player). As already anticipated, we mostly consider two preference types here: the \emph{objective type} maps each game
onto its own, true evolutionary payoffs (i.e., $ \theta = \pi_i $); the \emph{regret type} maps each game onto its negative regret
transformation (i.e., $ \theta = -\text{preg} $). We also consider two belief types: the
\emph{precise type} represents uncertainty about the co-players in a single probability measure
$\mu \in \Delta(\Acts_{-i})$; the \emph{imprecise type}
represents uncertainty as a set of probability measures $\Gamma \subseteq \Delta(\Acts_{-i})$.
It is convenient to conceptualize precise beliefs as a special case of imprecise beliefs by
looking at $\mu$ as a stand-in for the singleton set $\set{\mu}$.
The action selection function $\kappa_i(\gamma_i(G)) \subseteq A_i$ that we consider here is \emph{maximinimization of expected utility} (see Appendix~\ref{sec:impr-prob-beli} for motivations).

@@

Evidence from experimental literature suggests that agents are mostly uncertainty averse
\textcolor{red}{[REF]}. In line with empirical data, we assume that players in the population
are uncertainty averse and conform to maxmin expected utility. The resulting behavior is then
produced by maximinimizing the subjective utility given by the player's preference type over
the set of probabilities given by the belief type.

>>>>

\documentclass[fleqn,reqno,12pt]{article}

@@

\usepackage{bigints}
\usepackage{MnSymbol}

\usepackage{geometry}
 % \geometry{
 % a4paper,
 % left=300mm,
 % right=250mm,
 % top=400mm,
 % bottom=400mm
 % }

\pagenumbering{gobble} 
\usepackage[document]{ragged2e}

\doublespacing

@@

\title{Smart Representations: {R}ationality and Evolution in a Richer  Environment}

\date{}

@@

\begin{abstract}
  Standard applications of evolutionary game theory look at a single, fixed game and focus on
  the evolution of behavior for that game alone. Instead, this paper uses tools from
  evolutionary game theory to study the evolutionary competition between \emph{choice
    mechanisms} in a rich and variable environment. A choice mechanism is a way of representing
  a decision situation, paired with a method for choosing an act based on this 
  subjective representation. We demonstrate the usefulness of this approach by a case study
  that compares a number of classic proposals for decision making. Our
  main result is that regret-based preference representations can be evolutionarily selected
  for, despite being non-veridical.
\end{abstract}

@@

If agents deal with a rich and variable environment, they have to face many different choice
situations.
Standard evolutionary game models frequently
simplify reality in two ways that are relevant for present purposes. Firstly, the environment
is represented as a fixed \emph{stage game}; secondly, the focus of
evolutionary selection is behavior for that stage game alone. In contrast, some argue for
studying the evolutionary competition of general \emph{choice mechanisms} in a rich and
variable environment
\citep[e.g.,][]{FawcettHamblin2013:Exposing-the-be,McNamara2013:Towards-a-Riche,HammStev12}. To show that
there is no tension here, this paper introduces a method for reconciling established notions of
evolutionary game theory with evolutionary selection of choice mechanisms in variable
environments \citep[see, inter alia, related work
by][]{Harley1981:Learning-the-Ev,ZollmanSmead2010:Plasticity-and-,SmeadZollman2013:The-Stability-o,OConnor2016:Evolving-to-Gen}.

@@

The subjective representation scheme $\gamma_i$ of agent $i$ is a function that
takes a game $G = \tuple{ N, (A_i , \pi_i)_{i \in N} }$ and maps it onto
$\gamma_i(G) = (\theta, e)$ where
$\theta \mycolon \bigtimes_{j \in N} A_j \rightarrow \mathbb{R}$ is a subjective utility function (the \emph{preference type} of the player), and $e$ is some subjective representation of a player's uncertainty about the co-players'
behavior (the \emph{epistemic} or \emph{belief type} of the player). As already anticipated, we mostly consider two preference types here: the \emph{objective type} maps each game
onto its own, true evolutionary payoffs (i.e., $ \theta = \pi_i $); the \emph{regret type} maps each game onto its negative regret
transformation (i.e., $ \theta = -\text{preg} $). We also consider two belief types: the
\emph{precise type} represents uncertainty about the co-players in a single probability measure
$\mu \in \Delta(\Acts_{-i})$; the \emph{imprecise type}
represents uncertainty as a set of probability measures $\Gamma \subseteq \Delta(\Acts_{-i})$.
It is convenient to conceptualize precise beliefs as a special case of imprecise beliefs by
looking at $\mu$ as a stand-in for the singleton set $\set{\mu}$.
The action selection function $\kappa_i(\gamma_i(G)) \subseteq A_i$ that we consider here is \emph{maximinimization of expected utility} (see Appendix~\ref{sec:impr-prob-beli} for motivations).

@@

Evidence from experimental literature suggests that agents are mostly uncertainty averse \citep[e.g.,][]{TrautKuil16}. In line with empirical data, we assume that players in the population
are uncertainty averse and conform to maxmin expected utility. The resulting behavior is then
produced by maximinimizing the subjective utility given by the player's preference type over
the set of probabilities given by the belief type.
